---
title: "MATH1318 Time-Series Analysis Assignment 1"
output:
  word_document: default
  html_notebook: default
---
### Milind Shaileshkumar Parvatia
### s3806853

# Time Series Analysis of thickness of Ozone


```{r}
rm(list = ls())
library(readr)
library(xts)
library(TSA)
```


```{r}

dataset <- read.csv("data1.csv", header = FALSE)

rownames(dataset) <- seq(from=1927, to=2016)

timeseries <- ts(dataset$V1, start = 1927, end = 2016, frequency = 1)

plot(timeseries,ylab='Thickness level in dobson units', xlab='Year',type='o', main = "Time Series plot of thickness of Ozone layer 1927 to 2016")

pacf(dataset, main = "PACF plot of thickness of Ozone layer")
acf(dataset, main = "ACF plot of thickness of Ozone layer")


```
From the fig 1 we can discuss following descriptive points from seires.


```{r}
plot(y=timeseries, x=zlag(timeseries), ylab='Thickness level in dobson units',xlab='Previous Year Thickness level', main = "Scatter Plot of thickness of Ozone layer in consequtive years")

y = timeseries              # Read the color data into y
x = zlag(timeseries)        # Generate  of the color series
index = 2:length(x)    # Create an index to get rid of the first NA value in x
cor(y[index],x[index]) # Calculate correlation between numerical values in x and y```

```
```{r}
# Created function to implement model plotting everytime
modelPlotting <- function(timeseries, model, modelname){
  
  #Timeseries plot of fitted model
  plot(timeseries,
     ylim = c(min(c(fitted(model), as.vector(timeseries))),
              max(c(fitted(model), as.vector(timeseries)))), 
     ylab='Dobson units',
     xlab='Years',
     type='o',
     main = "Thickness of Ozone layer 1927 to 2016")
  lines(ts(fitted(model), start = 1927))
  
  par(mfrow=c(3,2))
  
  #Histogram plot of Standardized Residuals
  hist(rstudent(model),xlab='Standardized Residuals', main = c(paste0("Histogram of the standardized residuals from ", toString(modelname))))
  
  #Timeseries plot of residuals of model
  plot(y=rstudent(model),x=as.vector(time(timeseries)),xlab='Time', ylab='Standardized Residuals',type='o', main = c(paste0(toString(modelname), "'s plot of residuals with labels")))
  
  # Normal qq-plot of residuals of model
  y = rstudent(model)
  qqnorm(y, main = c(paste0("Normal Q-Q plot of the standardized residuals from ", toString(modelname))))
  qqline(y, col = 2, lwd = 1, lty = 2)

  # ACF plot
  acf(y, main = c(paste0("ACF plot of thickness of Ozone layer of ", toString(modelname))))
  
  # PACF plot
  pacf(y, main = c(paste0("PACF plot of thickness of Ozone layer of ", toString(modelname))))
  
  par(mfrow=c(1,1))
  
  #Shapiro Wilk test
  shapiro.test(y)
}

```

```{r}
# Linear Model 1 summary and plots

model1 = lm(timeseries ~ time(timeseries)) # label the model as model1
summary(model1)

modelPlotting(timeseries, model1, "The Linear model")

```

```{r}
# Quadratic Model 1 summary and plots

t = time(timeseries)
t2 = t^2
model2 = lm(timeseries~t+t2) # label the model as model1
summary(model2)

modelPlotting(timeseries, model2, "The Quadratic model")

```
```{r}
# Seasonal Model 1 summary and plots

timeseries1 <- ts(dataset$V1, start = 1927, frequency = 10)

year.=season(timeseries1) # period added to improve table display and this line sets up indicators
model3 = lm(timeseries1~ year.-1) # -1 removes the intercept term
summary(model3)

modelPlotting(timeseries, model3, "The Seasonal model")

```


```{r}
# Harmonic Model 4 summary and plots

har.=harmonic(timeseries, 0.4) # calculate cos(2*pi*t) and sin(2*pi*t)
model4.har = lm(timeseries ~ har.-1)
summary(model4.har)

modelPlotting(timeseries, model4.har, "The Harmonic model")

```
```{r}


h = 5 # 5 steps ahed forecasts
# Now we will implement the two-step algorithm

t.start = time(timeseries)[1]
t.end = time(timeseries)[length(timeseries)]

t = seq((t.end+1), (t.end+h), 1)
t2 = t^2

modelT = model2 # label the model as model1

new = data.frame(t,t2)

forecasts = predict(modelT, new, interval = "prediction")
print(forecasts)


mergedForcastsSeries <- c(timeseries, forecasts[,1])

plot(timeseries,
     ylim = c(min(c(as.vector(mergedForcastsSeries)-5)), max(c(as.vector(mergedForcastsSeries)+5))),
     xlim = c(t.start, t.end+5),
     ylab = "Dobson units", 
     xlab='Years',
     type = 'o',
     main = "Thickness of Ozone layer 1927 to 2016")
# We need to convert forecasts to time series object starting from the first 
# time steps-ahead to be able to use plot function. 
# We do this for all columns of forecasts
lines(ts(as.vector(forecasts[,1]), start = 2017), col="red", type="l")
lines(ts(as.vector(forecasts[,2]), start = 2017), col="blue", type="l")
lines(ts(as.vector(forecasts[,3]), start = 2017), col="blue", type="l")
legend("topleft", lty=1, pch=1, col=c("black","blue","red"), text.width = 22,c("Data","5% forecast limits", "Forecasts"))
 

```

# Task 2


```{r}

acf(timeseries , main = "ACF plot of thickness of Ozone layer")
pacf(timeseries , main = "PACF plot of thickness of Ozone layer")

```


```{r}

positiveData <- timeseries + abs(min(timeseries)) +1
plot(positiveData,
     ylab = "Dobson units", 
     xlab='Years',
     type = 'o',
     main = "Thickness of Ozone layer 1927 to 2016")

OzoneLog <- log(positiveData)
plot(OzoneLog,
     ylab = "Dobson units", 
     xlab='Years',
     type = 'o',
     main = "Thickness of Ozone layer 1927 to 2016")

OzoneLogDiff <- diff(log(positiveData), differences = 1)
plot(OzoneLogDiff,
     ylab = "Dobson units", 
     xlab='Years',
     type = 'o',
     main = "Thickness of Ozone layer 1927 to 2016")

acf(OzoneLogDiff , main = "ACF plot of thickness of Ozone layer")
pacf(OzoneLogDiff , main = "PACF plot of thickness of Ozone layer")

```
```{r}
eacf(diff(log(abc)))
```

```{r}
eacf(diff(log(abc), differences = 1))
```


```{r}

BICModel = armasubsets(y=diff(log(abc), differences = 1), nar=6, nma=6, y.name='test', ar.method='ols')
plot(BICModel)

```